A fully labelled proof system for intuitionistic modal logics

### 2020-09-29 FullyLabelled/JLC/review-answer.txt

One of the main concerns of Reviewer 1 is that we have not cited all
the relevant literature. We thank the reviewer for pointing out the
missing references and we added them accordingly.

We now address the other comments:

>  Perhaps even more importantly, the authors of the paper under
>  review neglect an important caveat in the design of rules of
>  contraction-free labelled sequent calculi: to guarantee that
>  contraction be admissible one needs to ensure that a certain
>  closure condition be satisfied

We mentioned in the submitted version that contraction is derivable 
in our system from the monotonicity rules, which in turn
are admissible. We made this clearer in the new version in
Remark 3.4.

@@

>  In Section 7, if the rules G_klmn are added without their
>  contracted instances, the calculus will not be complete. For
>  example, in the classical case, Euclidean transitivity ∀a, b,
>  c(aRb&aRc → bRc) has the contracted instance ∀a, b(aRb → bRb) that
>  allows for the derivation of BOX(BOX A → A). This, however, can
>  hardly be derived without the rule’s contracted instance.

In fact, the contracted instances are added. In the same way as for
other rules, occurring labels can be equal and substituted as needed. 
We failed to make this explicit in the first submitted version and 
corrected that mistake in Remark 7.2.

@@

>  On top of this, the authors’ claim that labelled calculi ”work best
>  for logics with standard Kripke semantics” is utterly wrong:
>  labelled calculi have been developed for several variants of Kripke
>  semantics, such as Lewis’s semantics for counterfactuals,
>  neighbourhood semantics, Kripke semantics of impossible worlds, and
>  so on.

We meant to say that for logics with standard Kripke semantics 
there is this close relationship between the semantics and the labels in 
the proof system which makes this method both efficient and elegant.

@@

>  In conclusion, the paper does not look highly original. The
>  birelational semantics is not a novelty, nor a calculus
>  internalising that semantics appears an original idea.

We do not claim the birelational semantics to be a novelty of this work
(We acknowledge that it dates back to the 1980s). Our paper is using the 
semantical results of Plotkin and Sterling and others to design proof 
systems for extensions of intuitionistic modal logic that had not been 
studied before. We try to make this clearer in the new version (see Section 7).

@@

>  Also the reason for preferring one semantics for intuitionistic
>  modal logic over another should be discussed in detail, as
>  otherwise the authors’ work would rather look a routine exercise

We chose to work the standard birelational semantics for intuitionistic modal 
logic IK. There are indeed other possible semantics that have been studied 
for other constructive interpretations of modal logics such as birelational 
semantics with impossible words or neighbourhood semantics. These could 
potentially be studied in a similar framework to ours and we commented on 
this topic in the conclusion.

@@

On the other side, Reviewer 2 finds this paper well written, but has some 
technical concerns. We address them below.

>  The main issue is indeed with the "wide range of logics”. While the
>  base logic is discussed in detail, the range of logics is treated
>  in an “Extensions” section that I believe could actually be more
>  detailed.

We put more emphasis on Section 7 in this new version. We hope we made 
it clearer how to generate "stronger logics" adding new axioms to our system. 
We extended this section with more explanation and also added a few comments 
and remarks. For example, as a better way to understand the rules and the 
completeness proof, we present the specific case of the 4-axiom.

@@

>  Most importantly, I am not convinced by the gklmn rule. The side
>  condition stipulates that y’ and u are fresh, whereas the frame
>  property contains existential quantifications. I was thus expecting
>  either a Skolemization in the style of Vigano in the “Labelled
>  Non-Classical Logics” book, or a geometric rule with existential
>  quantification in the style of Simpson [Sim94] or, perhaps even
>  better, a “mixed” rule where the labels are assumed more or less in
>  the style of an \exists elimination to derive a labeled formula.  I
>  think that a thorough discussion of the alternatives, along with a
>  more thorough proof of the correctness of the chosen solution, is
>  needed.

We hope the additional Remark 7.4 answers these concerns. 
We are indeed using the same method that Simpson (or Negri) applied to 
geometric formulas, but in our case also allowing the preorder atoms to occur. 
This method enforces that both relational and preorder atoms only appear on 
the left-hand-side of a sequent, never on the right-hand-side. 
As suggested, it should also be possible to treat the Scott-Lemmon axioms in 
the style of Viganò and atoms would then appear on the right of sequents 
only and indeed Skolemisation of some of the quantifiers would be required.

@@

>  By the way, it does not help at all that the applications of the
>  rule g1111 are merged with \Box_L and \Diamond_R in the example
>  derivation.

Following this comment, we fixed the derivation which should now 
show better the use of rule g1111.

@@

>  Also Remark 4.2 could benefit from some additional explanation. You
>  write that the proof of Theorem 4.1 shows the “need” of the rules
>  F_1, refl and trans. However, strictly speaking, the proof only
>  shows how the rules are used and that they are helpful to prove the
>  axioms. To show the “need" you would likely have to show that no
>  other proofs of the axioms are possible. That is of course the
>  case, but unless you prove it formally, it is not fully correct to
>  claim that the “need” is shown. This is pedantic, I know.

In this case, our intention was indeed to only show how to use those rules and how 
critical there are in these specific cases.

>Reconsider the abstract. It reads very operational and does not clearly state the
>contributions of the paper.

We rewrote the abstract added some more context and with the aim of better reflecting 
our contribution in the paper.

We want to thank the reviewers again for their time and for helping us improve our paper. 
We hope to have addressed all their insightful comments both in this letter and in the new version of this paper.

Best regards,
Sonia Marin, Marianela Morales and Lutz Strassburger

### 2020-09-29 FullyLabelled/JLC/review-answer.txt

Following this comment, we fixed the derivation which should now 
show better the use of rule g1111.

>  Also Remark 4.2 could benefit from some additional explanation. You
>  write that the proof of Theorem 4.1 shows the “need” of the rules
>  F_1, refl and trans. However, strictly speaking, the proof only
>  shows how the rules are used and that they are helpful to prove the
>  axioms. To show the “need" you would likely have to show that no
>  other proofs of the axioms are possible. That is of course the
>  case, but unless you prove it formally, it is not fully correct to
>  claim that the “need” is shown.

In this case, our intention was indeed to only show how to use those rules and how 
critical there are in these specific cases.

> Finally, I suggest that the authors reconsider the abstract, which, 
> in my opinion, does not do their paper justice. It reads very “operational” 
> and does not clearly state the contributions of the paper.

We rewrote the abstract adding some more context and with the aim of better reflecting 
our contributions

### 2020-09-29 FullyLabelled/JLC/conclusion.tex
concl

\review{Also the reason for preferring one semantics for intuitionistic modal logic over another should be discussed in detail, as otherwise the authors’ work would rather look a routine exercise.}
\todo{add a paragraph in conclusion: saying that it is an interesting idea to apply method to CK, but extension is tricky \cite{ADS:lmcs}} \sonia{done?}
We have considered in this work the logic known as $\IK$ (and its extensions) with respect to the birelational semantics that is its most well-studied semantics.
%
As we have mentioned, however, different basis can be considered for intuitionistic modal logic, for example the one called \emph{constructive modal logic} $\mathsf{CK}$ and any in between.
%
These can also be studied within the birelational framework with some additional conditions, and we are convinces the fully labelled framework, once extended with these conditions, will be suitable to treat constructive fragments equally.
%
However, treating extensions of those with axioms such as we did here for Scott-Lemmon logics is known to be complex~\cite{ADS:lmcs}.

>>>>

\review{Also the reason for preferring one semantics for intuitionistic modal logic over another should be discussed in detail, as otherwise the authors’ work would rather look a routine exercise.}
\todo{add a paragraph in conclusion: saying that it is an interesting idea to apply method to CK, but extension is tricky \cite{ADS:lmcs}} \sonia{done?}
We have considered in this work the logic known as $\IK$ (and its extensions) with respect to the birelational semantics that is its most well-studied semantics.
%
As we have mentioned, however, different basis can be considered for intuitionistic modal logic, for example the one called \emph{constructive modal logic} $\mathsf{CK}$~\cite{xxx} and any in between.
%
These can also be studied within the birelational framework with some additional conditions, and we are convinced that the fully labelled framework, once extended with these conditions, will be suitable to treat constructive fragments equally.
%
However, treating extensions of $\mathsf{CK}$ with axioms such as we did here for Scott-Lemmon logics is known to be complex~\cite{arisaka:etal:lmcs15}.
